---
title: "Data Management Group Assignment"
output: html_document
date: "2024-03-07"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment=NA,attr.source='.numberLines')

rm(list=ls())
#install.packages("readr")
#install.packages("RSQLite")
#install.packages("dplyr")
install.packages("chron")
install.packages("ggplot2")
library(readr)
library(RSQLite)
library(dplyr)
library(chron)
library(ggplot2)
```

```{r connect}
my_connection <- RSQLite::dbConnect(RSQLite::SQLite(),"e-commerce.db")
                            
```

Create Category table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Category(
  category_id VARCHAR(20) PRIMARY KEY NOT NULL,
  category_name VARCHAR (20) NOT NULL,
  parent_id INT,
  parent_name VARCHAR (20)
  );
```

Create Customer table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Customer(
  customer_id VARCHAR(50) PRIMARY KEY NOT NULL,
  email VARCHAR (100) NOT NULL,
  first_name VARCHAR (100) NOT NULL,
  last_name VARCHAR (100) NOT NULL,
  street_name VARCHAR (100) NOT NULL,
  post_code VARCHAR(64) NOT NULL,
  city VARCHAR (100) NOT NULL,
  password_c VARCHAR (10) NOT NULL, 
  phone_number INT (11) NOT NULL,
  referral_by VARCHAR(20)
  );
```

Create Supplier table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Supplier (
    seller_id VARCHAR(50) PRIMARY KEY NOT NULL,
    seller_store_name VARCHAR(100),
    supplier_email VARCHAR(255),
    password_s VARCHAR(255),
    receiving_bank VARCHAR(50),
    seller_rating INT,
    seller_phone_number VARCHAR(20),
    seller_address_street VARCHAR(255),
    s_post_code VARCHAR(50),
    s_city VARCHAR(50)
    );

```

Create Warehouse table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Warehouse (
    warehouse_id VARCHAR(50) PRIMARY KEY NOT NULL,
    capacity INT,
    current_stock INT,
    w_city VARCHAR(50),
    w_post_code VARCHAR(50),
    w_address_street VARCHAR(255)
    );
```

Create Product table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Product (
  product_id INT PRIMARY KEY NOT NULL,
  product_name VARCHAR(50),
  category_id VARCHAR(20),
  warehouse_id VARCHAR(50),
  seller_id VARCHAR(50),
  product_weight FLOAT,
  product_price FLOAT,
  FOREIGN KEY (seller_id) REFERENCES Supplier(seller_id)
  FOREIGN KEY (category_id) REFERENCES Category(category_id),
  FOREIGN KEY (warehouse_id) REFERENCES Warehouse(warehouse_id)
  );
```

Create Shipment table

```{sql connection=my_connection}
CREATE TABLE IF NOT EXISTS Shipment (
    shipment_id VARCHAR(50) PRIMARY KEY NOT NULL,
    shipping_method VARCHAR(50),
    shipping_charge FLOAT
    );
```

Create Orders table

```{sql connection=my_connection}
 CREATE TABLE IF NOT EXISTS Orders (
    order_id VARCHAR(50) NOT NULL,
    order_date DATE NOT NULL,
    order_status VARCHAR(50),
    quantity_of_product_ordered INT,
    payment_method VARCHAR(50),
    voucher_value INT,
    review_rating INT,
    shipment_id VARCHAR(50),
    product_id VARCHAR(50) NOT NULL,
    customer_id VARCHAR(50) NOT NULL,
    PRIMARY KEY (order_id, customer_id, product_id),

    FOREIGN KEY (shipment_id) REFERENCES Shipment(shipment_id),
    FOREIGN KEY (customer_id) REFERENCES Customer(customer_id),
    FOREIGN KEY (product_id) REFERENCES Product(product_id)
    );
```

Read .csv files in

```{r dataloading,message=FALSE,warning=FALSE}
Category <- readr::read_csv("category_data.csv")
Category$category_id <- as.character(Category$category_id)
Category$parent_id <- as.character(Category$parent_id)


Customer <- readr::read_csv("customer_data.csv")
Customer$customer_id <- as.character(Customer$customer_id)


Supplier <- readr::read_csv("supplier_data.csv")
Supplier$seller_id <- as.character(Supplier$seller_id)


Warehouse <- readr::read_csv("warehouse_data.csv")
Warehouse$warehouse_id <- as.character(Warehouse$warehouse_id)


Product <- readr::read_csv("product_data.csv")
Product$product_id <- as.character(Product$product_id)
Product$seller_id <- as.character(Product$seller_id)
Product$warehouse_id <- as.character(Product$warehouse_id)
Product$category_id <- as.character(Product$category_id)


Shipment <- readr::read_csv("shipment_data.csv")
Shipment$shipment_id <- as.character(Shipment$shipment_id)


Orders <- readr::read_csv("order_data.csv")

Orders$order_date <- as.Date(Orders$order_date, format = "%Y/%m/%d")
Orders$order_date <- as.character(Orders$order_date)
Orders$order_id <- as.character(Orders$order_id)
Orders$customer_id <- as.character(Orders$customer_id)
Orders$product_id <- as.character(Orders$product_id)
Orders$shipment_id <- as.character(Orders$shipment_id)

#db exucute query

```

Write them to the database

```{r writebacktodb}
#RSQLite::dbWriteTable(my_connection,"Category",Category,overwrite=TRUE)

#RSQLite::dbWriteTable(my_connection,"Category",Category,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Customer",Customer,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Supplier",Supplier,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Warehouse",Warehouse,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Product",Product,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Shipment",Shipment,append=TRUE)
#RSQLite::dbWriteTable(my_connection,"Orders",Orders,append=TRUE)

```

Use this to verify the table we created

```{sql connection=my_connection}
PRAGMA table_info(Orders);
```

```{sql connection=my_connection}
SELECT * FROM Orders
LIMIT 10

```
```{sql connection=my_connection}
SELECT * FROM Category
LIMIT 10

```
Data analysis and visualization

Rank order value from highest to lowest

```{sql connection=my_connection}
SELECT 
    o.order_id,
    o.customer_id,
    SUM(o.quantity_of_product_ordered * (p.product_price - o.voucher_value)) AS total_value
FROM 
    Orders o
JOIN 
    Product p ON o.product_id = p.product_id
GROUP BY 
    o.order_id
ORDER BY 
    total_value DESC;
```

Warehouse capacity v.s. current stock level

```{r barplot}
barplot(Warehouse$capacity, col = "steelblue", ylim = c(0, max(Warehouse$capacity, Warehouse$current_stock)),
        main = "Warehouse Capacity and Current Stock", xlab = "Warehouse ID", ylab = "Quantity")
barplot(Warehouse$current_stock, col = "lightpink", add = TRUE)
legend("topright", legend = c("Capacity", "Current Stock"), fill = c("steelblue", "lightpink"))
```

Product price distribution

```{r}
# Calculate the mean price
mean_price <- mean(Product$product_price)

# Create the histogram
ggplot(Product, aes(x = product_price)) +
  geom_histogram(binwidth = 1, position = "identity") +
  geom_vline(xintercept = mean_price, linetype = "dotted", color = "darkred") +  # Add the mean line
  labs(x = "Product Price", y = "Frequency", fill = "Category ID",
       title = "Distribution of Product Prices by Category") +
  theme_minimal()

```

Number of customers in each city

```{r}
city_counts <- Customer %>%
  group_by(city) %>%
  summarise(num_customers = n())

# Plot the counts
ggplot(city_counts, aes(x = reorder(city, -num_customers), y = num_customers, fill = city)) +
  geom_bar(stat = "identity") +
  labs(x = "City", y = "Number of Customers",
       title = "Number of Customers in Each City") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Product review rating ranking from highest to lowest

```{r}
# Calculate the average rating for each product
class(Orders$review_rating)
Orders$review_rating <- as.numeric(Orders$review_rating)
product_ratings <- Orders %>%
  group_by(product_id) %>%
  summarise(avg_rating = mean(review_rating, na.rm = TRUE))

# Sort products by average rating in descending order
product_ratings <- product_ratings[order(-product_ratings$avg_rating),]

top_products <- product_ratings[product_ratings$avg_rating == 5,]


ggplot(product_ratings, aes(x = reorder(product_id, -avg_rating), y = avg_rating, fill = factor(product_id %in% top_products$product_id))) +
  geom_bar(stat = "identity") +
  labs(x = "Product ID", y = "Average Rating",
       title = "Average Rating for Each Product") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0)) +
  scale_fill_manual(values = c("grey80", "darkred"), guide = FALSE)

```
